Large and Infinitary Quotient Inductive-Inductive Types

### 2020-01-07 draft/paper.tex
paper

\todo{univ lvl for pi sigma}

%% Both for functions and $\Sigma$, the output universe level is given as the
%% maximum of the levels of the constituent type.

We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
have equality reflection and uniqueness of identity proofs (UIP). The unit type
is $\top$, with inhabitant $\tt$.

@@

We use cumulative universes and cumulative subtyping as described
in \cite{timany2018cumulative}. Concretely, we have a $\blank\leq\blank$ subtyping
relation on types, and a corresponding subtyping rule.
\begin{mathpar}
  \inferrule*{i \leq j}
             {\Gamma \vdash \Set_i \leq \Set_j}

@@

We need to talk about constructions at arbitrary universe levels. For the sake
of simplicity, we do not assume a notion of universe polymorphism in cETT,
instead we quantify over levels in an unspecified theory outside of cETT. Hence,
a universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family
of cETT terms. This is sufficient in this paper, since we do not need to
internalize level polymorphism.

We reuse the notation of cETT functions for
universe polymorphism, e.g.\ as in the following function:
\[
\lambda\,i.\,\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
\]

@@

\begin{definition}[Notion of model for the theory of signatures]\label{def:tos}
For levels $i$ and $j$, $\ToS_{i,j} : \Set_{\max(i,\,1+j)}$ is a cETT type whose
elements are ToS models (or ToS-algebras). $\ToS_{i,j}$ is an iterated
$\Sigma$-type, containing all of the following components.

@@

\emph{Notational conventions.} We name elements of $\Con$ as $\Gamma$, $\Delta$,
$\Theta$, elements of $\Sub\,\Gamma\,\Delta$ as $\sigma$, $\delta$, $\nu$,
elements of $\Ty\,\Gamma$ as $A$, $B$, $C$, and elements of $\Tm\,\Gamma\,A$ as
$t$, $u$, $v$. CwF components by default support de Bruijn indices, which are
not easily readable. We use instead a nameful notation for binders in context
extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon\,\ext (a : U) \ext (t :
\El\,a))$. We also define a type-theoretic flavor of $\appi$ for convenience:
\begin{alignat*}{3}
  & \blank\appitt\blank && :
      \Tm\,\Gamma\,(\Pii\,a\,B)\ra
      (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[\id, u])\\
  & t\appitt u && :\equiv (\appi\,t)[\id, u]
\end{alignat*}
We abbreviate non-dependent inductive $\Pii$ as $\blank\arri\blank$, and likewise we use $\blank\arre\blank$ and $\blank\arrinf\blank$.

@@

\begin{definition}[Notion of signature]\label{def:signature}
A QIIT signature at level $i$ is a context in an arbitrary p$M : \ToS_{i,j}$
model. We define the type of such signatures as follows:
\[
  \Sig_i :\equiv (j : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
\]

@@

In the following, for $i$ and $j$ levels we define a model $\bM_{i,j} : \ToS_{i,
  1+\max(i,1+j)}$, such than $\Con_{\bM_{i,j}}$ is a type of structured categories (of
algebras). The level $i$ marks the level of all external sets in function
domains, and $j$ marks the level of all internal sorts in an algebra. Hence,
every algebra has level $\max(i,1+j)$. The bump is only needed for $j$, since
algebras only contain elements of $T : \Set_j$ types, while inductive sets are
themselves elements of $\Set_i$. For example, $NatAlg_i : \Set_{\max(0, 1+i)} :
\Set_{1+\max(0, 1+i)}$,

@@

We define $\bCon : \Set_{1+\max(i,1+j)}$ as $\flCwF_{\max(i,1+j)}$.
%% We define
%% $\flCwF$ as follows.

@@

\todo{segue text?}

@@

\todo{we have a level lift here, decide how to handle}

%% We define $\boldsymbol{\Sub\,\Gamma\,\Delta} : \Set_{1 + \max(i, j+1)}$ as the type
%% of weak flCwF morphisms from $\boldsymbol{\Gamma}$ to $\boldsymbol{\Delta}$.

>>>>

Both for functions and $\Sigma$, the output universe level is given as the
maximum of the levels of the constituent types, e.g.\ $(x : A)\ra B : \Set_{\max(i,j)}$
when $A : \Set_i$ and $B : \Set_j$.

We write propositional equality as $t = u$, with $\refl_t$ for reflexivity. We
have equality reflection and uniqueness of identity proofs (UIP). The unit type
is $\top : \Set_0$, with inhabitant $\tt$.

@@

We use cumulative universes and cumulative subtyping as described
in \cite{timany2018cumulative}. Concretely, we have a $\blank\leq\blank$ subtyping
relation on types, specified by the following rules:
\begin{mathpar}
  \inferrule*{i \leq j}
             {\Gamma \vdash \Set_i \leq \Set_j}

@@

We need to talk about constructions at arbitrary universe levels. For the sake
of simplicity, we do not assume a notion of universe polymorphism in cETT,
instead we quantify over levels in an unspecified theory outside of cETT. Hence,
a universe polymorphic cETT term is understood as a $\mathbb{N}$-indexed family
of cETT terms. We reuse the notation of cETT functions for
universe polymorphism, e.g.\ as in the following function:
\[
\lambda\,i.\,\Set_i : (i : \mathbb{N})\ra \Set_{i+1}
\]

@@

\begin{definition}[Notion of model for the theory of signatures]\label{def:tos}
For levels $i$ and $j$, $\ToS_{i,j} : \Set_{\max(i+1,\,j+1)}$ is a cETT type whose
elements are ToS models (or ToS-algebras). $\ToS_{i,j}$ is an iterated
$\Sigma$-type, containing all of the following components.

@@

\emph{Notational conventions.} We name elements of $\Con$ as $\Gamma$, $\Delta$,
$\Theta$, elements of $\Sub\,\Gamma\,\Delta$ as $\sigma$, $\delta$, $\nu$,
elements of $\Ty\,\Gamma$ as $A$, $B$, $C$, and elements of $\Tm\,\Gamma\,A$ as
$t$, $u$, $v$. CwF components by default support de Bruijn indices, which are
not easily readable. We use instead a nameful notation for binders in context
extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon\,\ext (a : U) \ext (t :
\El\,a))$. We also define a type-theoretic flavor of $\appi$ for convenience:
\begin{alignat*}{3}
  & \blank\appitt\blank && :
      \Tm\,\Gamma\,(\Pii\,a\,B)\ra
      (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[\id, u])\\
  & t\appitt u && :\equiv (\appi\,t)[\id, u]
\end{alignat*}
We abbreviate non-dependent inductive $\Pii$ as $\blank\arri\blank$, and
likewise we use $\blank\arre\blank$ and $\blank\arrinf\blank$ for non-dependent
external and infinitary functions.

@@

\begin{definition}[Notion of signature]\label{def:signature}
A QIIT signature at level $i$ is a context in an arbitrary $M : \ToS_{i,j}$
model. We define the type of such signatures as follows:
\[
  \Sig_i :\equiv (j : \mathbb{N})\ra(M : \ToS_{i,j})\ra \Con_M
\]

@@

In the following, for $i$ and $j$ levels we define a model $\bM_{i,j} : \ToS_{i,
  \max(i,j+1)+1}$, such than $\Con_{\bM_{i,j}}$ is a type of structured categories (of
algebras). The level $i$ marks the level of all external sets in function
domains, and $j$ marks the level of all internal sorts in an algebra. Hence,
every algebra has level $\max(i,j+1)$. The bump is only needed for $j$, since
algebras merely contain elements of $A : \Set_i$ types, while inductive sets are
themselves elements of $\Set_j$. For example, $NatAlg_j : \Set_{\max(0, j+1)} :
\Set_{\max(0, j+1)+1}$,

@@

We define $\bCon : \Set_{\max(i,j+1)+1}$ as $\flCwF_{\max(i,j+1)}$.
%% We define
%% $\flCwF$ as follows.

@@

In order to talk about weak structure-preservation in the interpretation of substitutions, we need to specify isomorphisms for contexts and types.

@@
